Formalizing the Edmonds-Karp Algorithm

### 2016-05-23 paper/root.tex
---

  During this project, we have learned some lessons on verified algorithm development: 
  \begin{itemize}
  \item It is important to keep the levels of abstraction strictly separated.
    For example, when implementing the capacity function with arrays, one needs to show that it is only applied to valid nodes.
    However, proving that, e.g., augmenting paths only contain valid nodes is hard at this low level. 
    Instead, one can protect the application of the capacity function by an assertion--- already on a high abstraction level where it can be easily discharged. 
    On refinement, this assertion is passed down, and ultimately available for the implementation.
    Optimally, one wraps the function together with an assertion of its precondition into a new constant, which is then refined independently.
  \item Profiling has helped a lot in identifying candidates for optimization. For example, based on profiling data, we decided to delay a 
    possible deforestation optimization on augmenting paths, and to first refine the algorithm to operate on residual graphs directly.
  \item ``Efficiency bugs'' are as easy to introduce as for unverified software. For example, out of convenience, we implemented the successor list computation by
    \emph{filter}. Profiling then indicated a hot-spot on this function. As the order of successors does not matter, we invested a bit more work to make the computation tail 
    recursive and gained a significant speed-up. Moreover, we realized only lately that we had accidentally implemented and verified
    matrices with column major ordering, which have a poor cache locality for our algorithm. Changing the order resulted in another significant speed-up.
  \end{itemize}

>>>>

  During this project, we have learned some lessons on verified algorithm development: 
  \begin{itemize}
  \item It is important to keep the levels of abstraction strictly separated.
    For example, when implementing the capacity function with arrays, one needs to show that it is only applied to valid nodes.
    However, proving that, e.g., augmenting paths only contain valid nodes is hard at this low level. 
    Instead, one can protect the application of the capacity function by an assertion --- already on a high abstraction level where it can be easily discharged. 
    On refinement, this assertion is passed down, and ultimately available for the implementation.
    Optimally, one wraps the function together with an assertion of its precondition into a new constant, which is then refined independently.
  \item Profiling has helped a lot in identifying candidates for optimization. For example, based on profiling data, we decided to delay a 
    possible deforestation optimization on augmenting paths, and to first refine the algorithm to operate on residual graphs directly.
  \item ``Efficiency bugs'' are as easy to introduce as for unverified software. For example, out of convenience, we implemented the successor list computation by
    \emph{filter}. Profiling then indicated a hot-spot on this function. As the order of successors does not matter, we invested a bit more work to make the computation tail 
    recursive and gained a significant speed-up. Moreover, we realized only lately that we had accidentally implemented and verified
    matrices with column major ordering, which have a poor cache locality for our algorithm. Changing the order resulted in another significant speed-up.
  \end{itemize}
